---
name: Luigui
surname: Gallardo-Becerra
position: "Software Engineer & Data Scientist"
phone: +1 619 602 0725
www: www.lmgb.xyz
email: "luiguimichelgallardo@gmail.com"

github: LuiguiGallardo
linkedin: luiguigallardo
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Software Engineer with 6+ years of experience in designing scalable data systems, automating complex workflows, and building cloud-native solutions. Proficient in Python, SQL, C#, with a strong background in backend development."
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
library(dplyr)
```
# Experience

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Graduate Research Assistant (Data Science & Bioinformatics)", "January 2019 - July 2025", "Institute of Biotechnology, UNAM",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```
* Built an internal AMP research data portal using ASP.NET Core and clean architecture, with SQL Server and Dapper for querying and visualizing large-scale peptide prediction datasets.
* Engineered automated, containerized pipelines using Nextflow, Docker, and Snakemake for reproducible data processing.
* Optimized distributed computing systems and high-performance computing (HPC) workflows for large-scale analysis.
* Developed machine learning models to analyze complex datasets, resulting in 14 peer-reviewed publications.
* Conducted statistical analyses and feature engineering to identify trends and patterns in biological data.

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Data Engineer", "October 2016 - December 2018", "Appen",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```

* Built robust SQL-based ETL architectures for data integration and automated reporting systems.
* Developed and automated data collection workflows using Python and Bash.
* Preprocessed and analyzed large-scale datasets for machine learning projects.

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Software Engineer (Internship)", "January 2016 - July 2016", "Dept. of Computer Science, CUCEI",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```

* Developed a cloud-deployed web application using ASP.NET Core for patient record management, contributing to both backend and frontend development.

# Education
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Ph.D. in Computational Biology", "January 2019 - July 2025", "National Autonomous University of Mexico (UNAM)", "Mexico City, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Master of Science in Computational Biology", "August 2016 - January 2019", "National Autonomous University of Mexico (UNAM)", "Mexico City, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor of Science in Molecular Biology", "August 2012 - January 2016", "University of Guadalajara (UDG)", "Guadalajara, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```



# Key Skills
* **Programming & Frameworks**: C#, ASP.NET Core, Python, SQL, Bash, JavaScript, React, R
* **Cloud & Infrastructure**: AWS, Docker, CI/CD, Linux, Database Management, HPC
* **Tools**: Git, Jupyter, Snakemake, Nextflow
* **Data Engineering**: ETL Pipelines, Apache Spark, System Automation, Large-Scale Data Processing
* **Soft Skills**: Problem-solving, Technical Leadership, Project Management, Agile Collaboration
* **Languages**: English (Full professional proficiency), Spanish (native)