---
name: Luigui
surname: Gallardo-Becerra
position: "Software Engineer & Bioinformatician"
phone: +1 619 602 0725
www: luiguigallardo.github.io/web_page/
email: "luiguimichelgallardo@gmail.com"

github: LuiguiGallardo
linkedin: luiguigallardo
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Experienced in software engineering, bioinformatics, and data science. Proven track record in developing robust software solutions, managing research projects, and contributing to scientific publications."
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
library(dplyr)
```
# Experience

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Graduate Research Assistant", "January 2019 - Current", "Institute of Biotechnology, UNAM",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```
* Developed software tools and pipelines to streamline data analysis and processing.
* Administered the laboratory's server, ensuring the installation of necessary software and maintaining secure server environments.
* Created reproducible analyses and reports, contributing to efficient project management.
* Presented research findings at various scientific seminars and conferences, effectively disseminating critical insights and findings to diverse audiences.

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Data Engineer", "October 2016 - December 2018", "Appen",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```

* Conducted data collection and preprocessing for various clients, including browsers and social networks.
* Developed and delivered weekly data deliverables, and collaborated on database creation tasks.

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Software Engineer (Internship)", "January 2016 - July 2016", "Dept. of Computer Science, CUCEI",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```

* Developed a cloud-deployed web application using ASP.NET Core for patient record management, contributing to both backend and frontend development.

# Education
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Ph.D. (Computational Biology)", "January 2019 - Current", "National Autonomous University of Mexico", "Mexico City, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Master of Science (Computational Biology)", "August 2016 - January 2019", "National Autonomous University of Mexico", "Mexico City, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor of Science (Molecular Biology)", "August 2012 - January 2016", "University of Guadalajara", "Guadalajara, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


# Key Skills
* **Programming Languages**: Python, C#, R, SQL, Bash, HTML, CSS, JavaScript, TypeScript
* **Tools & Frameworks**: HPC, Linux, Docker, Tableau, Nextflow, Snakemake, Jupyter Notebooks, Rstudio, Django, ASP.NET, React, Angular, Shiny
* **Software Development**: Web application development, database management, version control (Git), unit testing, Agile methodologies
* **Bioinformatics**: 16S rRNA profiling, Genome/Transcriptome assembly, Metagenomics, Metatranscriptomics, Viromics
* **Soft Skills**: Team player, fast learner, excellent communication, problem-solving, multi-tasking
* **Languages**: English (Full professional proficiency), and Spanish (native)
