---
name: Luigui
surname: Gallardo-Becerra
position: "Data Scientist & Bioinformatician"
phone: +1 619 602 0725
www: luiguigallardo.github.io/web_page/
email: "luiguimichelgallardo@gmail.com"

github: LuiguiGallardo
linkedin: luiguigallardo
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Data scientist and bioinformatician looking for new challenges to apply and expand my skills in a new company."
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
library(dplyr)
```

# About me
 * As a graduate researcher, I participated in different projects using modern Denomic technologies and Data Science applied to biological data to obtain novel information. Recently finished Google's "Data Analytics Professional Certificate", and Codecademy's "Data Scientist: Machine Learning Specialization".

# Experience
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Ph.D. Researcher", "2016-2022", "Institute of Biotechnology, National Autonomous University of Mexico",
) %>% 
  detailed_entries(Degree, Year, Institution)
```

I was part of different research projects. From the beginning, obtaining or creating the raw data, to the final report or publication. The tasks that I carried out includes the following:

* Manage research projects: the creation of main and secondary objectives/research questions.
* Server management (HPC): installation of Linux operative system, installation of specific software, avoiding server slowdowns, and building secure server environments.
* Creation of pipelines: using bash, python, Snakemake, and Rabix Composer, creating pipelines to avoid repetitive workflows and make reproducible analysis.
* Software development: using bash, python, Perl, and R, developing several software tools to solve and achieve specific goals.
* Creation of plots: using R, Rstudio, Jupyter Notebooks, Excel, Graphpad, and Tableau.
* Maintenance of GitHub account: the creation and maintenance of repositories on GitHub.
* Creation of reports/publication: using Markdown and Word, I delivered weekly and biannual reports to the members of the Laboratory. Also, I co-authored peer-reviewed publications that are highly cited in their respective areas.
* Presentations to the general public and academic events: participate in scientific dissemination, specialized seminars, and congress.

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Mathematics Teacher", "2014-2016", "Student Coaching, Guadalajara, Mexico",
) %>% 
  detailed_entries(Degree, Year, Institution)
```

* I taught math classes to middle and high school students to increase their grades for college admission exams.

# Education
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Ph.D.", "2019-2022", "Institute of Biotechnology, National Autonomous University of Mexico", "Cuernavaca, Mexico",
  "Master of Science", "2016-2019", "Institute of Biotechnology, National Autonomous University of Mexico", "Cuernavaca, Mexico",
  "Bachelor of Biology", "2012-2016", "University of Guadalajara", "Guadalajara, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Soft skills
* Team player who can also work independently with excellent communication skills
* Fast learner and could quickly adapt to new technologies.
* Excellent problem-solving capabilities.
* Can manage a significant workload or several projects from different areas without losing track of them.

# Programing Languages, Tools & Bioinformatic skills:
* Python, R, SQL, Bash, Perl, C++, LaTeX, HTML & CSS, JavaScript, High Performance Computing (HPC), Google Cloud, AWS, Linux, Tableau, Docker, Snakemake, Jupyter Notebooks, Rstudio, Django
* 16S rRNA profiling, Genome/Transcriptome assembly and annotation, Metagenomics, Metatranscriptomics, Viromics

# Languages
* English – Full professional proficiency
* Spanish – Native
