---
name: Luigui
surname: Gallardo-Becerra
position: "Data Scientist & Bioinformatician"
#address: "School of Physics & Chemistry, École Normale Supérieure"
phone: +1 619 602 0725
#profilepic: cv_photo.jpeg
www: luiguigallardo.github.io/web_page/
email: "luiguimichelgallardo@gmail.com"

github: LuiguiGallardo
linkedin: luiguigallardo
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Data scientist and bioinformatician looking for new challenges to apply and expand my skills in a new company."
output: vitae::moderncv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
library(dplyr)
```

# About me
 * As a graduate researcher, I participated in different projects using modern genomic technologies and Data Science applied to biological data to obtain novel information.
 * Recently finished Google's "Data Analytics Professional Certificate", and Codecademy's "Data Scientist: Machine Learning Specialization".

# Experience
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Ph.D. Researcher", "2016-2022", "Institute of Biotechnology, National Autonomous University of Mexico",
) %>% 
  detailed_entries(Degree, Year, Institution)
```

I was part of different research projects. From the beginning, obtaining or creating the raw data, to the final report or publication. The tasks that I carried out includes the following:
\break

* Manage research projects: the creation of main and secondary objectives/research questions.
* Server management (HPC): installation of Linux operative system, installation of specific software, avoiding server slowdowns, and building secure server environments.
* Creation of pipelines: using bash, python, Snakemake, and Rabix Composer, creating pipelines to avoid repetitive workflows and make reproducible analysis.
* Software development: using bash, python, Perl, and R, developing several software tools to solve and achieve specific goals.
* Creation of plots: using R, Rstudio, Jupyter Notebooks, Excel, Graphpad, and Tableau.
* Maintenance of GitHub account: the creation and maintenance of repositories on GitHub.
* Creation of reports/publication: using Markdown and Word, I delivered weekly and biannual reports to the members of the Laboratory. Also, I co-authored peer-reviewed publications that are highly cited in their respective areas.
* Presentations to the general public and academic events: participate in scientific dissemination, specialized seminars, and congress.

\

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Mathematics Teacher", "2014-2016", "Student Coaching, Guadalajara, Mexico",
) %>% 
  detailed_entries(Degree, Year, Institution)
```

* I taught math classes to middle and high school students to increase their grades for college admission exams.

# Education
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Ph.D.", "2019-2022", "Institute of Biotechnology, National Autonomous University of Mexico", "Cuernavaca, Mexico",
  "Master of Science", "2016-2019", "Institute of Biotechnology, National Autonomous University of Mexico", "Cuernavaca, Mexico",
  "Bachelor of Biology", "2012-2016", "University of Guadalajara", "Guadalajara, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Skills
* I am a fast learner and could quickly adapt to new technologies.
* If I don't know something, I always search for a way to learn it.
* I possess excellent problem-solving capabilities.
* I can manage a significant workload or several projects from different areas without losing track of them.

# Programing Languages:
* Python
* R
* SQL
* Bash
* Perl
* C++
* HTML & CSS

# Bioinformatics skills:
* 16S rRNA profiling
* Genome/Transcriptome assembly and annotation
* Metagenomics
* Metatranscriptomics
* Viromics

# Tools:
* Git
* High Performance Computing (HPC)
* Google Cloud
* AWS
* Linux
* Tableau
* Docker
* Snakemake
* Jupyter Notebooks
* Rstudio


# Languages
* English – Full professional proficiency
* Spanish – Native


# Publications
```{r}
bibliography_entries("packages.bib") %>%
  arrange(desc(author$family), issued)
```

# Certifications and Awards
* Google Data Analytics Specialization - 2022
* Data Scientist: Machine Learning Specialization, Codecademy - 2021
* Honorific mention in Bachelor Degree - 2017
* First place in the XVI State Contest of Physics Apparatus and Experiments - 2011
* Honorable mention in ExpoCiencias Nacional, Tlaxcala 2010 with science dissemination project - 2010
* Second place in the State Competition of the XXIII Mexican Mathematics Olympiad - 2010
* Third place in the XIX University Olympiad of Mathematics - 2009
* Second place in the State Competition of the XXI University Olympiad of Mathematics - 2009

