---
name: Luigui
surname: Gallardo-Becerra
position: "Data Scientist & Bioinformatician"
phone: +1 619 602 0725
www: www.lmgb.xyz
email: "luiguimichelgallardo@gmail.com"

github: LuiguiGallardo
linkedin: luiguigallardo
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Data Scientist with a strong background in bioinformatics, data analysis, and machine learning. Experienced in processing large-scale datasets, developing predictive models, and implementing data pipelines. Proficient in Python, R, and SQL with expertise in statistical analysis, and visualization."
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
library(dplyr)
```
# Experience

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Graduate Research Assistant", "January 2019 - Current", "Institute of Biotechnology, UNAM",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```
* Developed machine learning models to analyze metagenomic and transcriptomic datasets.
* Created reproducible bioinformatics pipelines using Nextflow and Snakemake.
* Managed and optimized high-performance computing (HPC) workflows.
* Conducted statistical analyses to identify biomarkers in complex biological datasets.

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Data Engineer", "October 2016 - December 2018", "Appen",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```

- Preprocessed and analyzed large-scale datasets for various AI-driven projects.
- Implemented SQL-based ETL pipelines for data integration.
- Automated data collection and preprocessing workflows using Python and Bash.

```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Software Engineer (Internship)", "January 2016 - July 2016", "Dept. of Computer Science, CUCEI",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```

* Developed a cloud-deployed web application using ASP.NET Core for patient record management, contributing to both backend and frontend development.

# Education
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Ph.D. (Computational Biology)", "January 2019 - Current", "National Autonomous University of Mexico", "Mexico City, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Master of Science (Computational Biology)", "August 2016 - January 2019", "National Autonomous University of Mexico", "Mexico City, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor of Science (Molecular Biology)", "August 2012 - January 2016", "University of Guadalajara", "Guadalajara, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


# Key Skills
* **Programming Languages**: Python, C#, R, SQL, Bash, HTML, CSS, JavaScript, TypeScript
* **Tools & Frameworks**: HPC, Linux, Docker, Tableau, Nextflow, Snakemake, Jupyter Notebooks, Rstudio, Django, ASP.NET, React, Angular, Shiny
* **Software Development**: Web application development, database management, version control (Git), unit testing, Agile methodologies
* **Bioinformatics**: 16S rRNA profiling, Genome/Transcriptome assembly, Metagenomics, Metatranscriptomics, Viromics
* **Soft Skills**: Team player, fast learner, excellent communication, problem-solving, multi-tasking
* **Languages**: English (Full professional proficiency), and Spanish (native)
