---
name: Luigui
surname: Gallardo-Becerra
position: "Bioinformatician | Molecular Biologist"
phone: +1 619 602 0725
www: luiguigallardo.github.io/web_page/
email: "luiguimichelgallardo@gmail.com"

github: LuiguiGallardo
linkedin: luiguigallardo
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Bioinformatician and Molecular Biologist looking for new challenges to apply and expand my skills. In recent years, I participated in different projects using modern Genomic technologies (NGS), Data Analysis, and Data Science, to obtain novel biological information that culminated in high-impact peer-reviewed scientific publications"
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
library(dplyr)
```

# Research experience
```{r}
tribble(
  ~ empty, ~ Degree, ~ Year, ~ Institution,
  "", "Graduate Research Assistant (Bioinformatics)", "January 2019 - Current", "Institute of Biotechnology, UNAM",  
) %>% 
  detailed_entries(empty, Year, Degree, Institution)
```

I was part of different research projects, from the beginning obtaining or creating the raw data to the final report or publication. The tasks that I carried out includes the following:

* Management of research projects: participated in the creation of main and secondary objectives and research questions from different projects to obtain novel biological information useful for the scientific community
* Server management (HPC): managed the Laboratory's server, installed Linux operative systems and specific software/packages, avoided server slowdowns, and built secure server environments
* Creation of pipelines: created pipelines to avoid repetitive workflows and make reproducible analysis using bash, Python, Nextflow, and Snakemake
* Software development: developed several software tools to solve and achieve specific goals using bash, Python, Perl, and R
* Creation of plots: using R, Rstudio, Jupyter Notebooks, Excel, and Graphpad
* Maintenance of version control (CI/CD): creation and maintenance of GitHub repositories and their continuous version control 
* Creation of reports/publications: using Markdown and Word, I delivered weekly and biannual reports to the members of the Laboratory. Also, I co-authored peer-reviewed publications that are highly cited in their respective areas
* Presentations to the general public and academic events: participated in scientific dissemination, specialized seminars, and congresses

# Education
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Ph.D. (Computational Biology)", "January 2019 - Current", "National Autonomous University of Mexico", "Mexico City, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Master of Science (Computational Biology)", "August 2016 - January 2019", "National Autonomous University of Mexico", "Mexico City, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor of Science (Molecular Biology)", "August 2012 - January 2016", "University of Guadalajara", "Guadalajara, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Teaching Experience
* Guest lecturer at Universidad de La Sabana (Chia, Colombia). December 2023.
* Microbiota Symposium from Theory to Clinical Practice, Colegio de Profesionales de la Nutrición de Querétaro y el Bajío. June 2021.
* Bioinformatics and Reproducible Research Workshop, Postgraduate Program in Biological Sciences, UNAM. October 2019.
* Guest lecturer at Center of Genomic Sciences, UNAM. August 2018.

# Directed Theses
* Bachelor's Thesis: Itzel Abigail Hernández Reyna, Bachelor's in Nutrition (Universidad Autónoma del Estado de Morelos), December 2019. "Impact of bacteriophages associated with childhood obesity on the intestinal metagenome."

# Soft Skills
* Effective team player with strong communication skills, capable of working independently.
* Quick learner with the ability to adapt to new technologies rapidly.
* Exceptional problem-solving abilities.
* Capacity to manage significant workloads or multiple projects across different areas while maintaining organization and efficiency.

# Programming Languages, Tools & Bioinformatic skills:
* Programming Languages: Python, C#, R, SQL, Bash, HTML & CSS, JavaScript, Typescript
* Web frameworks: Django, ASP.NET, Angular
* Other tools: High-Performance Computing (HPC), Linux, Tableau, Docker, Nextflow, Snakemake, Jupyter Notebooks, Rstudio, MySQL, PostgreSQL, MongoDB
* Bioinformatic skills: 16S rRNA profiling, Genome/Transcriptome assembly, and annotation, Metagenomics, Metatranscriptomics, Viromics

# Languages
* English – Full professional proficiency
* Spanish – Native

# Publications
1. Gallardo-Becerra, L., Cervantes-Echeverría, M., Cornejo-Granados, F., Vazquez-Morado, L. E., & Ochoa-Leyva, A. (2023). Perspectives in Searching Antimicrobial Peptides (AMPs) Produced by the Microbiota. Microbial Ecology, 87(1), 8. https://doi.org/10.1007/s00248-023-02313-8
2. Gallardo-Becerra, L., Cornejo-Granados, F., García-López, R., Valdez-Lara, A., Bikel, S., Canizales-Quinteros, S., López-Contreras, B. E., Mendoza-Vargas, A., Nielsen, H., & Ochoa-Leyva, A. (2020). Metatranscriptomic analysis to define the Secrebiome, and 16S rRNA profiling of the gut microbiome in obesity and metabolic syndrome of Mexican children. Microbial Cell Factories, 19(1), 61. https://doi.org/10.1186/s12934-020-01319-y
3. Bikel, S., Gallardo-Becerra, L., Cornejo-Granados, F., & Ochoa-Leyva, A. (2022). Protocol for the isolation, sequencing, and analysis of the gut phageome from human fecal samples. STAR Protocols, 3(1), 101170. https://doi.org/10.1016/j.xpro.2022.101170
4. Bikel, S., López-Leal, G., Cornejo-Granados, F., Gallardo-Becerra, L., García-López, R., Sánchez, F., Equihua-Medina, E., Ochoa-Romo, J. P., López-Contreras, B. E., Canizales-Quinteros, S., Hernández-Reyna, A., Mendoza-Vargas, A., & Ochoa-Leyva, A. (2021). Gut dsDNA virome shows diversity and richness alterations associated with childhood obesity and metabolic syndrome. iScience, 24(8), 102900. https://doi.org/10.1016/j.isci.2021.102900
5. Cervantes-Echeverría, M., Gallardo-Becerra, L., Cornejo-Granados, F., & Ochoa-Leyva, A. (2023). The Two-Faced Role of crAssphage Subfamilies in Obesity and Metabolic Syndrome: Between Good and Evil. Genes, 14(1), Article 1. https://doi.org/10.3390/genes14010139
6. Chino de la Cruz, C. M., Cornejo-Granados, F., Gallardo-Becerra, L., Rodríguez-Alegría, M. E., Ochoa-Leyva, A., & López Munguía, A. (2023). Complete genome sequence and characterization of a novel Enterococcus faecium with probiotic potential isolated from the gut of Litopenaeus vannamei. Microbial Genomics, 9(3), 000938. https://doi.org/10.1099/mgen.0.000938
7. Cornejo-Granados, F., Gallardo-Becerra, L., Leonardo-Reza, M., Ochoa-Romo, J. P., & Ochoa-Leyva, A. (2018). A meta-analysis reveals the environmental and host factors shaping the structure and function of the shrimp microbiota. PeerJ, 6, e5382. https://doi.org/10.7717/peerj.5382
8. Cornejo-Granados, F., Lopez-Zavala, A. A., Gallardo-Becerra, L., Mendoza-Vargas, A., Sánchez, F., Vichido, R., Brieba, L. G., Viana, M. T., Sotelo-Mundo, R. R., & Ochoa-Leyva, A. (2017). Microbiome of Pacific Whiteleg shrimp reveals differential bacterial community composition between Wild, Aquacultured and AHPND/EMS outbreak conditions. Scientific Reports, 7(1), 11783. https://doi.org/10.1038/s41598-017-11805-w
9. Jatuyosporn, T., Laohawutthichai, P., Romo, J. P. O., Gallardo-Becerra, L., Lopez, F. S., Tassanakajon, A., Ochoa-Leyva, A., & Krusong, K. (2023). White spot syndrome virus impact on the expression of immune genes and gut microbiome of black tiger shrimp Penaeus monodon. Scientific Reports, 13(1), Article 1. https://doi.org/10.1038/s41598-023-27906-8
10. Ochoa-Romo, J. P., Cornejo-Granados, F., Lopez-Zavala, A. A., Viana, M. T., Sánchez, F., Gallardo-Becerra, L., Luque-Villegas, M., Valdez-López, Y., Sotelo-Mundo, R. R., Cota-Huízar, A., López-Munguia, A., & Ochoa-Leyva, A. (2022). Agavin induces beneficial microbes in the shrimp microbiota under farming conditions. Scientific Reports, 12(1), Article 1. https://doi.org/10.1038/s41598-022-10442-2
11. Palomino-Hermosillo, Y. A., Berumen-Varela, G., Ochoa-Jiménez, V. A., Balois-Morales, R., Jiménez-Zurita, J. O., Bautista-Rosales, P. U., Martínez-González, M. E., López-Guzmán, G. G., Cortés-Cruz, M. A., Guzmán, L. F., Cornejo-Granados, F., Gallardo-Becerra, L., Ochoa-Leyva, A., & Alia-Tejacal, I. (2022). Transcriptome Analysis of Soursop (Annona muricata L.) Fruit under Postharvest Storage Identifies Genes Families Involved in Ripening. Plants, 11(14), 1798. https://doi.org/10.3390/plants11141798
