---
name: Luigui
surname: Gallardo-Becerra
position: "Data Scientist | Bioinformatician"
phone: +1 619 602 0725
www: luiguigallardo.github.io/web_page/
email: "luiguimichelgallardo@gmail.com"

github: LuiguiGallardo
linkedin: luiguigallardo
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Data Scientist and Bioinformatician passionate about Big Data, Computer Science, and Machine Learning, looking for new challenges to apply and expand my skills"
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
library(dplyr)
```

# About me
 * In recent years, I participated in different projects using modern Genomic technologies (NGS), Data Analysis, and Data Science, to obtain novel information that culminated in high-impact peer-reviewed scientific publications

# Experience
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Data Scientist | Bioinformatician", "January 2019 - December, 2022", "Institute of Biotechnology, UNAM",
) %>% 
  detailed_entries(Degree, Year, Institution)
```

I was part of different research projects, from the beginning obtaining or creating the raw data to the final report or publication. The tasks that I carried out includes the following:

* Management of research projects: participated in the creation of main and secondary objectives and research questions from different projects to obtain novel biological information useful for the scientific community
* Server management (HPC): managed the Laboratory's server, installed Linux operative systems and specific software/packages, avoided server slowdowns, and built secure server environments
* Creation of pipelines: created pipelines to avoid repetitive workflows and make reproducible analysis using bash, python, Snakemake, and Rabix Composer
* Software development: developed several software tools to solve and achieve specific goals using bash, python, Perl, and R
* Creation of plots: using R, Rstudio, Jupyter Notebooks, Excel, Graphpad, and Tableau
* Maintenance of GitHub account (CI/CD): creation and maintenance of GitHub repositories
* Creation of reports/publications: using Markdown and Word, I delivered weekly and biannual reports to the members of the Laboratory. Also, I co-authored peer-reviewed publications that are highly cited in their respective areas
* Presentations to the general public and academic events: participated in scientific dissemination, specialized seminars, and congresses

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Data Analyst (Part-time)", "October, 2016 - December, 2018", "Appen",
) %>% 
  detailed_entries(Degree, Year, Institution)
```
I was an independent contractor and participated in several projects that included Data Sourcing and Annotation for different Appen clients (browsers, social networks, etc.). The tasks I performed were the following:

* Data collection and Preprocessing: obtention of raw data and its preprocessing before passing it to the final client or the Machine Learning Department
* Creation of weekly deliverables: for the final client or project manager
* Content translation: from Spanish to English or English to Spanish

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Software Engineer (Internship)", "May, 2016 - July, 2016", "Department of Computer Science, CUCEI (UDG)",
) %>% 
  detailed_entries(Degree, Year, Institution)
```

I participated in the creation of Scientific Software Development. The tasks that I carried out included:

* Creation of an API: using .NET, I created an API to save patients' records and images

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Data Scientist | Bioinformatician (Internship)", "January, 2016 - July, 2016", "Institute of Biotechnology, UNAM",
) %>% 
  detailed_entries(Degree, Year, Institution)
```

I participated in the research project "Microbiome of Pacific Whiteleg shrimp reveals differential bacterial community composition between Wild, Aquacultured and AHPND/EMS outbreak conditions". The tasks that I performed were:

* Big Data Analysis: using a HPC, Linux and bash, I analyzed the 16S rRNA information of several samples
* Creation of plots: using R, Rstudio, Excel, and Matlab
* Software development: I developed several tools for ETL data using bash, python and Perl
* Creation of paper: with the data obtained from this internship, I co-autored this paper

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "Mathematics Teacher (Part-time)", "January, 2014 - December, 2015", "Student Coaching",
) %>% 
  detailed_entries(Degree, Year, Institution)
```
As a part-time teacher, I taught middle and high school students math classes to increase their grades or for college admission tests.

* Prepared group and individualized classes and lessons
* Applied weekly assessments to each student

# Education
```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Master of Science (Computer Science and Biochemistry)", "August, 2016 - January, 2019", "National Autonomous University of Mexico", "Mexico City, Mexico",
  "Bachelor of Science (Computer Science and Molecular Biology)", "August, 2012 - January, 2016", "University of Guadalajara", "Guadalajara, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Soft skills
* Team player who can also work independently with excellent communication skills
* Fast learner and could quickly adapt to new technologies.
* Excellent problem-solving capabilities.
* Can manage a significant workload or several projects from different areas without losing track of them.

# Programing Languages, Tools & Bioinformatic skills:
* Python, R, SQL, Bash, Perl, C#, LaTeX, HTML & CSS, JavaScript, High Performance Computing (HPC), Google Cloud, AWS, Linux, Tableau, Docker, Snakemake, Jupyter Notebooks, Rstudio, Django, ASP.NET, PostgreSQL, PySpark, MongoDB
* 16S rRNA profiling, Genome/Transcriptome assembly and annotation, Metagenomics, Metatranscriptomics, Viromics

# Languages
* English – Full professional proficiency
* Spanish – Native
